Updates a Procrustes and a PCA

### 2025-07-23 Analysis/17-update_procrustes_and_PCA.Rmd
Update 17-update_procrustes_and_PCA.Rmd

```{r, eval = FALSE, echo = FALSE}
## Get some random shapes for cross-validation

# Read in landmark data and landmark pairs file (necessary for enforcing symmetry later)
beaks <- readRDS("../Data/Raw/Beak_data/2020_06_19_MMB_LANDMARKS_ALL.rds")
pairs <- read.table("../Data/Raw/Beak_data/landpairs.txt")

@@

```{r, echo = FALSE}
## Loading the entire bird GPA
library(Morpho)
whole_gpa <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")

@@

```{r, echo = FALSE}
## But problem
## Good locations
plot(whole_gpa$PCscores[,c(1,2)], col = "grey", main = "Sanity check")
points(updated_pca[-c((dim(whole_gpa$PCscores)[1]+1): dim(updated_pca)[1]), c(1,2)], pch = 19, cex = 0.5, col = "grey")

@@

```{r, eval = FALSE, echo = FALSE}
orp <- function(A, mshape=NULL) { 
  p <- dim(A)[1]
  k <- dim(A)[2]
  n <- dim(A)[3]
  if (is.null(mshape))
      mshape <- arrMean3(A)

@@

## Centroid Sizing
mysize <- apply(new_landmarks,3,cSize)
for (i in 1:n){
    new_landmarks[,,i] <- new_landmarks[,,i]/mysize[i]
}

for (i in 1:n) {
    newdatarot[,,i] <- rotonto(x = whole_gpa$mshape, y =new_landmarks[,,i],scale=atts$scale,reflection=atts$reflect,centerweight=atts$centerweight,weights=atts$weights)$yrot
}

@@

## Getting the new PCAs
new_pca_orpda <- stats::predict(original_pca, newdata = proctan_ordat)#array.to(new_proc_tan, to = "matrix"))
new_pca_norop <- stats::predict(original_pca, newdata = proctan_norop)#array.to(new_proc_tan, to = "matrix"))

@@

>>>>

```r #, eval = FALSE, echo = FALSE}
## Get some random shapes for cross-validation
# Read in landmark data and landmark pairs file (necessary for enforcing symmetry later)
beaks <- readRDS("../Data/Raw/Beak_data/2020_06_19_MMB_LANDMARKS_ALL.rds")
pairs <- read.table("../Data/Raw/Beak_data/landpairs.txt")

@@

```r #, echo = FALSE}
## Loading the entire bird GPA
library(Morpho)
whole_gpa <- readRDS("../Data/Raw/Beak_data/2020_08_07_MMB_MORPHO_SHAPESPACE_FULL.rds")

@@

```r #, echo = FALSE}
## But problem
## Good locations
plot(whole_gpa$PCscores[,c(1,2)], col = "grey", main = "Sanity check")
points(updated_pca[-c((dim(whole_gpa$PCscores)[1]+1): dim(updated_pca)[1]), c(1,2)], pch = 19, cex = 0.5, col = "grey")

@@

```r #, eval = FALSE, echo = FALSE}
orp <- function(A, mshape=NULL) { 
  p <- dim(A)[1]
  k <- dim(A)[2]
  n <- dim(A)[3]
  if (is.null(mshape))
      mshape <- arrMean3(A)

@@

# Centroid Sizing -> skipped (cSizes all equal 1 already) -> maybe cSize needs to be a bit smaller? 0.95 seems to be the correct amount to get them to the right dimension.
mysize <- apply(new_landmarks,3,cSize)
for (i in 1:n){
    new_landmarks[,,i] <- new_landmarks[,,i]/mysize[i]
}

## The mean shape from the previously calculated gpa

## This value needs to be multiplied by 1.048-ish for some reason
prev_mean_shape <- whole_gpa$mshape # * 1.048

for (i in 1:n) {
    newdatarot[,,i]  <- rotonto(x = prev_mean_shape, y = new_landmarks[,,i], scale = atts$scale, reflection = atts$reflect, centerweight = atts$centerweight, weights = atts$weights)$yrot
}

@@

## Getting the new PCAs
new_pca_orpda <- stats::predict(original_pca, newdata = proctan_ordat)
new_pca_norop <- stats::predict(original_pca, newdata = proctan_norop)

@@

points(new_pca_orpda[, c(1,2)]/0.954, col = "blue", pch = 19, cex = 0.5)
points(new_pca_norop[, c(1,2)]*1.048, col = "orange", pch = 19, cex = 0.5)

## Something with 1.048 (or /0.95) scaling somewhere? Maybe cSize?
